REP DATA PEER ASSIGNMENT 1
========================================================

## Step 1: Loading and preprocessing the data

read in the data and set the 'Date' 
```{r echo = TRUE}
data <- read.csv("activity.csv", header = T)

```

## Step 2: What is mean total number of steps taken per day?


Make calculation of step set each day - and create array with the total step for each day

```{r}
steponeday <- tapply(data$steps, data$date, sum, na.rm = TRUE)
```

Make a histogram of the total number of steps taken each day

```{r fig.width=7, fig.height=6}
hist(steponeday, main ="Total number of step per day", xlab="steps")
```

### Calculate and report the mean and median total number of steps taken per day

First we reshape the data_set to be able to preform these calculations
```{r}
library(reshape2)
set_steponeday <- melt(data, id="date",measure.vars="steps")
steponeday <- dcast(set_steponeday, date ~ variable, sum, na.rm=TRUE)

```

The mean of the number of steps per day

```{r}
mean(steponeday[,2])
```

The median of the number of steps per day

```{r}
median(steponeday[,2])
```

## step 3: What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
data$interval <- as.factor(data$interval)

steponinterval_set <- melt(data, id="interval",measure.vars="steps")
steponinterval <- dcast(steponinterval_set, interval ~ variable, mean, na.rm = TRUE)
plot(steponinterval$interval,steponinterval$steps, type="l", main="Average step on time", xlab='time', ylab='frequency')
    
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max(steponinterval$steps)
subset(steponinterval, steponinterval$steps == max(steponinterval$steps))
```
We can see on 835h on average the most steps are made

## Step 4: Imputing missing values


Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
number.na <- is.na(data$steps)
table(number.na)
```
We find the total number of missing values is 2304

### Filling in missing Value in data set, using interval averages


By using for-loop to go to data set we update 'data' on missing values

```{r}
for (i in 1:nrow(data)){
  if (is.na(data$steps[i])){
    time <- data$interval[i]
    data$steps[i] <- subset(steponinterval$steps, steponinterval$interval == time)
  }
}

```

We do another check on the missing values and find there are no missing values
```{r}
number.na <- is.na(data$steps)
table(number.na)
```


Make calculation of step set each day - and create array with the total step for each day

```{r}
steponeday <- tapply(data$steps, data$date, sum, na.rm = TRUE)
```

Make a histogram of the total number of steps taken each day

```{r fig.width=7, fig.height=6}
hist(steponeday, main ="Total number of step per day", xlab="steps")
```

Calculate and report the mean and median total number of steps taken per day

First we reshape the data_set to be able to preform these calculations
```{r}
library(reshape2)
set_steponeday <- melt(data, id="date",measure.vars="steps")
steponeday <- dcast(set_steponeday, date ~ variable, sum, na.rm=TRUE)

```

The mean of the number of steps per day

```{r}
mean(steponeday[,2])
```

The median of the number of steps per day

```{r}
median(steponeday[,2])
```

### Conclusion

After filling the missing values the 'total' number of steps per day increased for all days. Just like the 'mean' and 'median' value.

## Step 5: difference in Weekdays and Weekend activities

First we set the date-variable as class(date) in the stepsonintval set
and we use weekdays-function to see the see which day it is.

```{r}
data$date <- as.Date(data$date)
day <- weekdays(data$date)
```

Function to set day name to Weekend and Weekday and add to data_set

```{r}
for (i in 1:length(day)){
  if (day[i] == "Saterday"){
    day[i] <- "Weekend"}
  else if (day[i] == "Sunday"){
    day[i] <- "Weekend"}
  else {day[i] <- "Weekday"}
}

data[,4] <- day
names(data)[4] <- "day"
```

Divide data in two sets"weekend" and "weekday"
Reshape new dataset for steponinterval for both sets

```{r}
weekday_set <- subset(data,data$day == "Weekday")
weekend_set <- subset(data, data$day == "Weekend")
steponinterval_set_wd <- melt(weekday_set, id="interval",measure.vars="steps")
steponinterval_wd <- dcast(steponinterval_set_wd, interval ~ variable, mean)
steponinterval_wd[,3] <- "Weekday"
names(steponinterval_wd)[3] <- "day"

steponinterval_set_we <- melt(weekend_set, id="interval",measure.vars="steps")
steponinterval_we <- dcast(steponinterval_set_we, interval ~ variable, mean)
steponinterval_we[,3] <- "Weekend"
names(steponinterval_we)[3] <- "day"

new_set <- rbind(steponinterval_we, steponinterval_wd)
```

plot the data
```{r}
library(lattice)
xyplot(steps ~ interval | day, data = new_set, type = 'l', layout = c(1,2))
```